---
title: "Regional Employment Indicators"
subtitle: "Quarterly statistical release"
author: "Bureau of Regional Statistics"
date: 2026-09-09
params:
  region: "Oregon"
format:
  html:
    template-partials:
      - title-block.html
---

```{r}
#| echo: false
#| output: asis
quarto::write_yaml_metadata_block(params = params)
```

```{r}
params$region
```


## Summary

The region is `r params$region`.
The same value is in the title block above, because the chunk at the top of
this file copies `params` into the document metadata.

## Rendering this report

With the defaults from the front matter:

```bash
quarto render report.qmd
```

The title block shows "Oregon". The summary above shows the same value.

With a new value on the command line:

```bash
quarto render report.qmd -P region:"California"
```

The title block shows "California". The summary above changes to match,
because the title block and the R code read the same `params`.

If you delete the chunk at the top of this file and render again with `-P`,
the title block goes back to "Oregon". Only the summary shows the new value.

## Detail

Body content stands in for the real release.
